The Young Planet DS Tuc Ab has a Low Obliquity\footnote{This paper includes data gathered with the 6.5 meter Magellan Telescopes located at Las Campanas Observatory, Chile.}

### 2019-10-31 paper/main.tex
paper draft complete

\newcommand{\todo}[3]{{\color{#2} \emph{#1} TO DO: #3}}
\newcommand{\mkntodo}[1]{\todo{ADINA}{cyan}{#1}}
\newcommand{\btmtodo}[1]{\todo{BEN}{blue}{#1}}
\newcommand{\anytodo}[1]{\todo{ANYONE}{green}{#1}}

@@

\newcommand{\carnegie}{Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena, CA 91101}

@@

\newcommand{\ames}{NASA Ames Research Center, Moffett Field, CA 94035, USA}

@@

\begin{document}
\title{The Young Planet DS Tuc Ab has a Low Obliquity}

@@

\author{Megan E. Bedell}
\affiliation{\flatiron}

@@

\author{Sharon Xuesong Wang}
\affiliation{\dtm}

@@

\author{Erin Flowers}
\affiliation{\princeton}

\author{Michael A. Gully-Santiago}
\affiliation{\ames}

@@

The abundance of short-period planetary systems with high obliquities is often taken as evidence that scattering processes play important roles in the formation and evolution of these systems.
More recent studies have suggested that wide binary companions can tilt protoplanetary disks, inducing a high obliquity on planets that form through smooth processes like disk migration.
\object{DS Tuc Ab}, a planet in the 40 Myr Tucana-Horologium association, likely traces its now-dissipated protoplanetary disk, enabling us to test these theories of disk physics. 
Here, we report on Rossiter-McLaughlin observations of one transit of DS Tuc Ab with the Planet Finder Spectrograph on the Magellan Clay Telescope at Las Campanas Observatory. 
We confirm the previously statistically validated planet by modeling the planet transit and stellar activity signals simultaneously, finding its projected obliquity to be low: $\psi = 12 \pm 13$ degrees.
This is the youngest planet to be observed using this technique; we provide a discussion on best practices to accurately measure the observed signal of similar young planets.
\end{abstract}

@@

Planet formation is poorly understood \citep[e.g.][]{Morbidelli16}. 
Each discovered planetary system represents an outcome of the planet formation process, and therefore provides an opportunity to learn about how different planets form in different environments. 
However, each observed present-day system is not a pure laboratory: over billions of years, planet-planet and planet-star gravitational interactions can scatter, torque, migrate, or otherwise perturb orbits, distancing planetary systems from their initial formation state \citep{Kozai62, Lidov62, Fabrycky07, Chatterjee08}. 
This picture is complicated by the fact that in many cases, stellar ages are very poorly known \citep[e.g.][]{Barnes07, Soderblom10}.

The origins of hot Jupiters are still unclear \citep{Dawson18}.
Kozai-Lidov cycles and tidal friction are often invoked to explain the formation of hot Jupiters \citep{Fabrycky07}, but smooth disk migration provides a reasonable alternative in many cases \citep{Ida08}. 
It is possible that multiple channels are required in order to explain all of the observed systems: \citet{Nelson17} analyze data from the HAT and WASP exoplanet surveys and find the data can be well-fit by a model in which $\sim85\%$ of hot Jupiters are formed through high-eccentricity migration and $\sim15\%$ through disk migration.

A population of low-obliquity planets is often considered a signature of smooth disk migration \citep{Morton11a, Ford14}. However, \citet{Albrecht12} show that obliquity is an imperfect tracer of the formation history for many stars, as the tidal realignment timescale for a massive, nearby planet can be shorter than the age of the system for many stars with convective outer layers.
\citet{Batygin12} suggest wide binary companions or nearby stars in the birth cluster can torque disks to random inclinations over the lifetime of the disk.

@@

Recently, data from the Transiting Exoplanet Survey Satellite \citep[\tess, ][]{Ricker14} were used to identify a planet with an orbital period of 8.14 days around the star \object{DS Tuc A} \citep{Benatti19, Newton19}. 
DS Tuc is a member of the Tucana-Horologium (Tuc-Hor) assocation, which has an age of 35-45 Myr \citep{Bell15, Crundall19}. 
DS Tuc itself is a binary with a projected separation of 240 AU. 
From \citet{Holman97}, the timescale for Kozai-Lidov interactions is
\begin{equation}
    \tau \approx P_\textrm{planet} \frac{M_\star}{M_\textrm{pert}} \bigg(\frac{a_\textrm{pert}}{a_\textrm{planet}}\bigg)^3 (1-e^2_\textrm{pert})^{3/2},
\label{eq:timescale}
\end{equation}
where $P_\textrm{planet}$ is the the orbital period of a planet with orbital semimajor axis  $a_\textrm{planet}$ about a host of mass $M_\star$, $M_\textrm{pert}$ is the
mass of the perturbing star, and $a_\textrm{pert}$ and $e_\textrm{pert}$ the semimajor axis and eccentricity of the outer object's orbit around the host star/planet system.

@@

As the age of the system is younger than this, even before considering the several million year timescale for planet formation, this planet likely traces the orientation of the now-dissipated protoplanetary disk. 
Measuring the obliquity of the planet relative to the star thus enables us to test theories of disk torqing.
We can measure the obliquity between the spin of the star and the orbit of the planet through the Rossiter-McLaughlin (R-M) effect, in which an apparent redshift and blueshift in the radial velocity of the star are observed as a transiting planet occults the blueshifted and redshifted hemispheres of the rotating star, respectively \citep{Rossiter24, McLaughlin24}.
DS Tuc Ab is the youngest known planet for which such an observation has been attempted, providing the best laboratory we have to test this theory.
We note that Zhou et al. (in prep) also obtained three transits of this planet, including two with the Planet Finder Spectrograph (PFS), for a complimentary Doppler tomographic analysis of this system.

@@

We obtained data with the PFS on the Magellan Clay Telescope \citep{Crane06, Crane08, Crane10}. 
On 2019 Aug 11 from UT 01:12 to UT 07:10 we obtained 49 measurements of the RV of DS Tuc A. 
The transit duration is 190 minutes, meaning approximately 50\% of the data were obtained in transit while the other half provide information about the out-of-transit RV baseline. Each exposure was 360 seconds in length and was taken with the 0\farcs3 $\times$ 2\farcs5 slit, which provides a resolution $R \approx 80,000$ per resolution element. All observations were taken with the iodine cell in place \citep{Marcy92}, which imprints a series of narrow lines at known wavelengths to measure the instrument point spread function and wavelength solution at each epoch.

\todo{pfs team}{red}{Can you please add a couple sentences here describing how Shk and Halpha measurements are calculated? Is the proper unit Angstroms for the derived values, as presented in Table 1?}

@@

We also collected a template spectrum of DS Tuc A on the night of the transit, immediately after the observations described above.
This spectrum, obtained under similar conditions and with the same slit but without the iodine cell in the light path, is used in the pipeline RV modeling.
All observations were then analyzed using the standard PFS pipeline \citep{Butler96}, which divides the spectrum into 2\AA\ chunks and fits each chunk independently.
The resultant RV is the weighted mean of the RVs of each chunk, and the provided uncertainty is the standard deviation of these velocities.

\todo{pfs team}{red}{Is there anything else you would like us to include?}

In fitting the template to the data for the night of the transit, we tested a reduction using all available data and one only using data from the night of the transit, finding a lower point-to-point scatter with the latter strategy. 
This is likely due to changes in the line profile shape on rotational timescales as the stellar surface varies, which leads to a redefinition in the median behavior of each 2\AA\ chunk.
As additional points are added at different stellar activity levels, this median behavior becomes more and more discrepant from that observed on the night of the transit, leading to a noisier fit. 
As a result, we use two different reductions when considering the data from the night of the transit itself and from the later observations, which typically have approximately twice the RV uncertainity despite generally being observed to the same expected SNR.
This strategy means in practice, there may be an RV offset between the two sections of the data set, but as we only consider the data from the transit night itself when modeling the R-M signal, this should not affect our results.

@@

\begin{figure}[!tbh]
  \begin{center}
    \includegraphics[width=0.5\textwidth, trim={0cm 0.0cm 0cm 0cm}, clip=true]{../figures/all_data.pdf}
   \end{center}
  \caption{RV time series for (left) the observed transit and (right) a timespan covering approximately one stellar rotation period. Note the different vertical scalings on each subplot. The R-M signal is easily detectable, occurring over a significantly different timescale than the rotationally-induced variability signal.}
  \label{fig:data}
\end{figure}

@@

We model the inferred radial velocities with version 1.0.0.dev3 of the \textit{starry} package of \citet{Luger19}. 
\textit{starry} models the star as a series of spherical harmonics. 
\todo{RODRIGO}{red}{please put a couple sentences about starry here, especially in how it does the R-M signal behind the scenes.}

@@

While the R-M signal model is always developed with \textit{starry} in our analysis, we test multiple approaches to model the effects of the star on these observations.
Over the six hours of the transit, the apparent radial velocity of the star increases by more than 50 m s$^{-1}$. 
In Section \ref{sec:trend} we argue this trend is due to stellar activity rather than an additional unseen planet.
Regardless of the cause of this signal, in order to accurately measure the obliquity of the transiting planet we must model the underlying stellar behavior as well.
We test several different approaches to this problem to ensure our results are not sensitive to our assumptions about the star.
We use low-order polynomials, up to the third degree, to fit the relatively long-term variability during the night of the transit. 
We also build a stellar activity model by fitting a function that is a linear sum of the observed $H\alpha$ and calcium $S_{HK}$ measurements during the night, which are correlated with the observed RV.
Finally, we also fit \textit{starry} models of a single starspot in time. The spot is modeled as a Gaussian darker region on the surface of the star; we allow the spot's size, contrast, and location to vary as we fit both signals.

@@

We test multiple likelihood functions with the same form.
First, we assume each data point is drawn from a mixture model which is the sum of two Gaussian functions, each with a different variance.
Under this model, each data point has some probability $q$ of being a ``good'' data point drawn from a relatively narrow distribution. This distribution is a combination of the PFS pipeline uncertainty and an additional jitter term, added in quadrature. 
Each data point also has a probability $1-q$ of being drawn from a much broader distribution, with a separate jitter term, to account for the possibility of stellar effects that significantly affect only a small number of data points. We allow the width of both Gaussian distributions to vary in our fitting procedure. 
Additionally, $q$ is a free parameter: if the data were well-modeled by a single Gaussian, we would find the posterior distribution on $q$ to be consistent with 1.0.
Forcing $q=1$ would have the same effect, fitting the data using only a single Gaussian.

@@

Our full likelihood function is then
\begin{equation}
\begin{split}
\mathcal{L} = \prod_i \bigg[ & \frac{q}{\sqrt{2\pi (\sigma^2 + s_1^2)}} \exp\bigg(-\frac{(y_{p,i} + y_{s,i}-V_i)^2}{2(\sigma^2 + s_1^2)}\bigg)  \\+ & \frac{1-q}{\sqrt{2\pi s_2^2)}} \exp\bigg(-\frac{(\alpha y_{p,i} + y_{s,i}-V_i)^2}{2(s_2^2)}\bigg) \bigg],
\end{split}
\end{equation}
where $i$ iterates over all observations, $\sigma$ the instrument uncertainties, $s_1$ and $s_2$ the additional jitter terms, $V$ the measured RV values, $y_p$ the component of the measured velocity from the planetary R-M signal, and $y_s$ the stellar activity model signal. $\alpha$ and $q$ retain their definitions from the previous paragraphs in this section.

@@

We fit these models to evaluate the posterior distribution using the \textit{emcee} package of \citet{Foreman-Mackey12}, an implementation of the affine-invariant ensemble sampler of \citet{Goodman10}.
For all explorations using the cubic polynomial out-of-transit model, we initialize 500 walkers; for all other runs we initialize 600 walkers. 
We run each for 3,000 steps, removing the first 2,000 as burn-in and considering the final 1,000 steps in our final analysis. 
We verify that our chains have converged following the method of \citet{Geweke92} and through visual inspection.

Here, we assume the planet orbit is circular. 
Given only information about the transit and assuming the orbital velocity of the planet does not change significantly during the transit there is a degeneracy between the eccentricity and the stellar radius. 
We choose to force circular orbits and fit the orbital separation $a/$\rstar\ as a free parameter. The opposite approach would be equally valid and produce similar results.
We include uninformative priors on all parameters except for the projected rotational velocity \vsini\ and the radius ratio $R_p$/\rstar.
For both parameters, \citet{Benatti19} and \citet{Newton19} provide discrepant results; we conservatively apply Gaussian priors with means of 18.3 km s$^{-1}$ and 0.057 and standard deviations of 1.8 km s$^{-1}$ and 0.003 for \vsini\ and $R_p$/\rstar, respectively.

@@

Our results are given in Table \ref{tab:results}. 
The maximum likelihood of the stellar activity indicator model is significantly lower than that of the other models. 
The maximum likelihood $\log \mathcal{L}_\textrm{max}$ differs by only 0.1 between the \textit{starry} and polynomial models, but is substantially lower for the third model. 
This corresponds to a Bayes' factor of $\approx 10^{-7}$ when comparing this model to the \textit{starry} model.
In a statistical sense, our starspot model provides approximately an equally valid fit to the data as our simple polynomial model; both provide a much more plausible fit than the stellar activity indicator model.

All models provide broadly consistent results on the measured obliquity. Considering the two families of most plausible models, the median of the obliquity posterior varies from 7 to 14 degrees depending on the specific choice of model used.
All statistical uncertainties range from 11 to 15 degrees.
Using our general polynomial fit, allowing both $q$ and $\alpha$ to vary, we infer an obliquity of $14 \pm 11$ degrees.
Likewise, with our \textit{starry} model we infer an obliquity of $12 \pm 13$ degrees.
As this model provides the highest likelihood fit to the data we choose this set of values as most representative of our knowledge of the obliquity of the system, but we emphasize that the particular choice of model or likelihood function does not appear to significantly affect the inferred obliquity.

\begin{figure}[!tbh]
  \begin{center}
    \includegraphics[width=0.45\textwidth, trim={0cm 0.0cm 1cm 1cm}, clip=true]{../figures/model_comp.pdf}
   \end{center}
  \caption{Draws from the posterior distributions of a simultaneous fit to the
  R-M signal and three different noise models. Green curves represent the noise model, while orange curves include the transit signal as well. From top to bottom, the noise models are a simple cubic polynomial fit, a fit regressed against the spectral stellar activity indicators, and a starspot model built with the \textit{starry} package of \citep{Luger19}. The polynomial and starspot models provide similar quality fits to the data, both of which find significantly higher log likelihood values than the third fit. Significantly, all give consistent results on the measured spin-orbit obliquity angle of approximately $12 \pm 12$ degrees. For visualization purposes, we interpolate the stellar activity indicator regression model between the observations with a cubic spline, although the fitting itself only requires this parameter to be defined at the times of the observations.}
  \label{fig:models}
\end{figure}

>>>>

\newcommand{\todo}[3]{{\color{#2} \emph{#1} TO DO: #3}}
\newcommand{\adftodo}[1]{\todo{ADINA}{cyan}{#1}}
\newcommand{\btmtodo}[1]{\todo{BEN}{blue}{#1}}
\newcommand{\anytodo}[1]{\todo{ANYONE}{green}{#1}}

@@

\newcommand{\carnegie}{The Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena, CA 91101, USA}

@@

\newcommand{\baeri}{Bay Area Environmental Research Institute, P.O. Box 25, Moffett Field, CA 94035, USA}

@@

\begin{document}
\title{The Young Planet DS Tuc Ab has a Low Obliquity\footnote{This paper includes data gathered with the 6.5 meter Magellan Telescopes located at Las Campanas Observatory, Chile.}}

@@

\author[0000-0001-9907-7742]{Megan E. Bedell}
\affiliation{\flatiron}

\author[0000-0002-4020-3457]{Michael A. Gully-Santiago}
\affiliation{\baeri}

@@

\author[0000-0002-6937-9034]{Sharon Xuesong Wang}
\affiliation{\carnegie}
\affiliation{\dtm}

@@

\author[0000-0001-8045-1765]{Erin Flowers}
\altaffiliation{\grfp}
\affiliation{\princeton}

@@

The abundance of short-period planetary systems with high obliquities is often taken as evidence that scattering processes play important roles in the formation and evolution of these systems.
More recent studies have suggested that wide binary companions can tilt protoplanetary disks, inducing a high obliquity on planets that form through smooth processes like disk migration.
\object{DS Tuc Ab}, a planet in the 40 Myr Tucana-Horologium association, likely traces its now-dissipated protoplanetary disk, enabling us to test these theories of disk physics. 
Here, we report on Rossiter-McLaughlin observations of one transit of DS Tuc Ab with the Planet Finder Spectrograph on the Magellan Clay Telescope at Las Campanas Observatory. 
We confirm the previously statistically validated planet by modeling the planet transit and stellar activity signals simultaneously. We find the planet's projected obliquity to be low: $\psi = 12 \pm 13$ degrees, suggesting that this planet likely formed through smooth disk processes and its protoplanetary disk was not significantly torqued by DS Tuc B.
This is the youngest planet to be observed using this technique; we provide a discussion on best practices to accurately measure the observed signal of similar young planets.
\end{abstract}

@@

Each discovered planetary system represents an outcome of the planet formation process, and therefore provides an opportunity to learn about how different planets form in different environments. 
However, each observed present-day system is not a pure laboratory: over billions of years, planet-planet and planet-star gravitational interactions can scatter, torque, migrate, or otherwise perturb orbits, distancing planetary systems from their initial formation state \citep{Kozai62, Lidov62, Fabrycky07, Chatterjee08}. 
This picture is complicated by the fact that in many cases, stellar ages are very poorly known \citep[e.g.][]{Barnes07, Soderblom10}.
These factors make it challening to develop and test models of planet formation which can explain all observations.

The origins of hot Jupiters are still unclear \citep{Dawson18}.
Kozai-Lidov cycles and tidal friction are often invoked to explain the formation of hot Jupiters \citep{Fabrycky07}, but smooth disk migration provides a reasonable alternative in many cases \citep{Ida08}. 
Multiple channels may be required to explain all of the observed systems: \citet{Nelson17} analyze data from the HAT and WASP exoplanet surveys and find the data can be well-fit by a model in which $\sim85\%$ of hot Jupiters are formed through high-eccentricity migration and $\sim15\%$ through disk migration.

A population of low-obliquity planets is often considered a signature of smooth disk migration \citep{Morton11a, Ford14}. However, \citet{Albrecht12} show that obliquity is an imperfect tracer of the formation history for many stars, as the tidal realignment timescale for a massive, nearby planet can be shorter than the age of the system for many stars with convective outer layers.
\citet{Batygin12} suggests wide binary companions or nearby stars in the birth cluster can torque disks to random inclinations over Myr timescales. In these cases, young planets in binary systems will have random obliquities even at ages of a few Myr, rather than these obliquities being excited by the companion over much slower timescales.

@@

Recently, Sector 1 data from the Transiting Exoplanet Survey Satellite \citep[\tess, ][]{Ricker14} were used to identify three transits of a planet with an orbital period of 8.14 days around the star \object{DS Tuc A} \citep{Benatti19, Newton19}. 
These papers statistically validated and characterized this planet, finding it to have a radius of $5.70 \pm 0.17$ \rearth and a transit duration of $0.1324 \pm 0.0005$ days. 
The host star has spectral type G6V, a stellar effective temperature of $5430 \pm 80$ K, and a model-dependent radius of $0.964 \pm 0.029$ or $0.872 \pm 0.027$ \rsun, from \citet{Newton19} and \citet{Benatti19}, respectively.
We refer the reader to those two papers for a detailed discussion of the stellar parameters and transit fits.

DS Tuc is a member of the Tucana-Horologium (Tuc-Hor) association, which has an age of 35-45 Myr \citep{Bell15, Crundall19}. 
DS Tuc itself is a binary, with a K3V companion at a projected separation of 240 AU. 
From \citet{Holman97}, the timescale for Kozai-Lidov interactions is
\begin{equation}
    \tau \approx P_\textrm{planet} \frac{M_\star}{M_\textrm{pert}} \bigg(\frac{a_\textrm{pert}}{a_\textrm{planet}}\bigg)^3 (1-e^2_\textrm{pert})^{3/2},
\label{eq:timescale}
\end{equation}
where $P_\textrm{planet}$ is the the orbital period of a planet with orbital semimajor axis  $a_\textrm{planet}$ about a host of mass $M_\star$, $M_\textrm{pert}$ is the
mass of the perturbing star, and $a_\textrm{pert}$ and $e_\textrm{pert}$ the semimajor axis and eccentricity of the outer object's orbit around the host star/planet system.

@@

As the age of the system is younger than this timescale, the planet has not had time to undergo these oscillations and is more likely to trace the orientation of the now-dissipated protoplanetary disk. 
Measuring the obliquity of the planet relative to the star thus enables us to test theories of disk torquing.
We can measure the projected obliquity between the spin of the star and the orbit of the planet through the Rossiter-McLaughlin (R-M) effect, in which an apparent redshift and blueshift in the radial velocity of the star are observed as a transiting planet occults the blueshifted and redshifted hemispheres of the rotating star, respectively \citep{Rossiter24, McLaughlin24}.
DS Tuc Ab is the youngest known planet for which such an observation has been attempted, providing the best laboratory we have to test this theory.
We note that Zhou et al. (in prep) also obtained three transits of this planet, including two with the Planet Finder Spectrograph (PFS), for a complementary Doppler tomographic analysis of this system. These two works use different data sets and analysis techniques, providing independent checks of the methods and assumptions made in each work.

@@

We obtained data with the PFS on the Magellan Clay Telescope \citep{Crane06, Crane08, Crane10}. 
On 2019 Aug 11 from UT 01:12 to UT 07:10 we obtained 49 measurements of the RV of DS Tuc A. 
The transit duration is 190 minutes, meaning approximately 50\% of the data were obtained in transit while the other half provides information about the out-of-transit RV baseline. Each exposure was 360 seconds in length and was taken with the 0\farcs3 $\times$ 2\farcs5 slit at 1$\times$2 binning, which provides a resolution of $R \approx 130,000$ at about 3 pixels per resolution element. All observations were taken with the iodine cell in place \citep{Marcy92}, which imprints a series of narrow lines at known wavelengths to measure the instrument point spread function and wavelength solution at each epoch.

From these spectra, we also derive two spectroscopic activity indicators: the emission flux measured in the Ca \textsc{II} H \& K lines, $S_{\rm HK}$, and the emission in the H$\alpha$ line, ``$S_{\rm H\alpha}$.'' $S_{\rm HK}$ is defined as in \cite{duncan1991} with the updated $R$ continuum area center from \cite{santos2000}, and $S_{\rm H\alpha}$ is defined as in \cite{gomesdasilva2011}.

@@

We also collected a template spectrum of DS Tuc A on the night of the transit, immediately after the observations described above.
This spectrum, obtained under similar conditions and with the same slit but without the iodine cell in the light path, is used in the pipeline RV modeling.
All observations were then analyzed using the standard PFS pipeline \citep{Butler96b}, which divides the spectrum into 2\AA\ chunks and fits each chunk independently.
The resultant RV is the weighted mean of the RVs of each chunk, and the provided uncertainty is the standard deviation of these velocities.

In our extraction of RVs from the available spectra, we tested both an extraction using all available data and one only using data from the night of the transit, finding a lower point-to-point scatter with the latter strategy. 
This is likely due to changes in the line profile shape on rotational timescales as the stellar surface varies, which leads to a redefinition in the median behavior of each 2\AA\ chunk.
The stellar template used in the extraction was obtained on the same night as the transit.
As additional points are added at different stellar activity levels, the median behavior of each chunk becomes more and more discrepant from that observed on the night of the transit, leading to a noisier fit. 
As a result, we use the results of two different extractions when considering the data from the night of the transit itself and from the later observations, which typically have approximately twice the RV uncertainty despite generally being observed to the same expected SNR.
This strategy means in practice, there may be an RV offset between the two subsections of the data, but as we only consider the data from the transit night itself when modeling the R-M signal, this should not affect our results.

@@

\begin{figure}[!tbh]
  \begin{center}
    \includegraphics[width=0.5\textwidth, trim={0cm 0.0cm 0cm 0cm}, clip=true]{../figures/all_data.pdf}
   \end{center}
  \caption{RV time series for (left) the observed transit and (right) a timespan covering approximately one stellar rotation period, as inferred from \tess\ photometry and discussed in Section \ref{sec:spots}. Note the different vertical scalings on each subplot. The R-M signal is easily detectable, occurring over a significantly different timescale than the rotationally-induced variability signal.}
  \label{fig:data}
\end{figure}

@@

We model the inferred radial velocities with version 1.0.0.dev3%
\footnote{\url{https://github.com/rodluger/starry/tree/dev}}%
of the \texttt{starry} package of \citet{Luger19}. The \texttt{starry} package models the stellar surface brightness and velocity field as an expansion of spherical harmonics, under which light curve and radial velocity computations are analytic. As the ability to model the Rossiter-McLaughlin effect is currently only available in the development version of the code, we describe it briefly here.

The radial velocity anomaly $\Delta\mathrm{RV}$ due to the Rossiter-McLaughlin may be expressed as \citep[e.g.][]{Gimenez06}
%
\begin{align}
    \label{eq:DeltaRV}
    \Delta\mathrm{RV} &= \frac{\int I(x, y) v(x, y) \mathrm{d}S}{\int I(x, y)\mathrm{d}S},
\end{align}
%
where $I(x,y)$ is the intensity at a point $(x, y)$ on the projected disk of the star, $v(x, y)$ is the radial velocity at that point, and the integrals are taken over the visible portion of the projected disk.
The radial component of the velocity field may be written in Cartesian coordinates as \citep[e.g.][]{Short18}
%
\begin{align}
\label{eq:f_xyz}
v(x, y) = \omega_{eq}(Ax + By)\left(1 - \alpha(-Bx + Ay + Cz)^2\right),
\end{align}
%
where $z=\sqrt{1 - x^2 - y^2}$, 
%
\begin{align}
    A &\equiv \sin(i)\cos(\psi) \nonumber \\
    B &\equiv \sin(i)\sin(\psi) \nonumber \\
    C &\equiv \cos(i),
\end{align}
%
and $i$ and $\psi$ are the stellar inclination and obliquity, respectively. 
The constant $\alpha$ is the linear shear due to differential rotation, whose
effect is to scale the rotational angular velocity of the star according to
%
\begin{align}
\omega(\theta) = \omega_{eq}(1 - \alpha \sin^2\theta),
\end{align}
%
where $\omega(\theta)$ is the angular rotational velocity as a function of the latitude $\theta$ and $\omega_{eq}$ is the angular velocity at the equator. 
The coordinate system adopted in Equation~(\ref{eq:f_xyz}) is such that the $x-y$ plane is the plane of the sky and $\hat{z}$ points toward the observer. The planet is assumed to transit the star along the $\hat{x}$ direction in a counterclockwise orbit about $\hat{y}$. The obliquity $\psi$ is therefore the component of the misalignment angle between the stellar rotation axis and the angular momentum vector of the planet projected onto the plane of the sky.

Because Equation~(\ref{eq:f_xyz}) is a third-degree polynomial in $x$, $y$, and $z$, it may be expressed exactly as a spherical harmonic expansion of degree $l=3$ \citep{Luger19}. This is done by expanding the products in Equation~(\ref{eq:f_xyz}) to obtain the coefficients multiplying each term in the polynomial basis \citep[Equation~7 in][]{Luger19} and transforming them via the spherical harmonic change of basis matrix \citep[Equation~9 in][]{Luger19}.%
\footnote{For a derivation, see \url{https://rodluger.github.io/starry/v1.0.0.dev5/notebooks/RossiterMcLaughlin.html}, which is also included as a notebook in the arXiv source material of this manuscript.}%

Under the assumption of quadratic limb darkening, the intensity field $I(x, y)$ may also be expressed exactly as a spherical harmonic expansion of degree $l=2$ \citep{Luger19}. Since the product of two spherical harmonics is also a spherical harmonic, the integrand in Equation~(\ref{eq:DeltaRV}) may be expressed exactly as a degree $l=5$ spherical harmonic expansion. We may therefore use the machinery in \texttt{starry} to \emph{analytically} compute both integrals in Equation~(\ref{eq:DeltaRV}) when the star is occulted by a planet. 
Note that a similar approach was used by \citet{Bedell19}, who model the Rossiter-McLaughlin effect of the hot Jupiter HD\,189733\,b with \texttt{starry}.

Finally, in the sections that follow we explore models in which the star has a single dark Gaussian-shaped spot. In \texttt{starry}, surface features such as spots are also modeled as an expansion of spherical harmonics up to a degree $l_\mathrm{max}$. Again, since spherical harmonics are closed under multiplication, the intensity field $I(x, y)$ of a quadratically limb-darkened spotted surface is simply a sum of spherical harmonics of degree $l_\mathrm{max} + 2$. The total degree of the velocity-weighted intensity (the integrand in Equation~\ref{eq:DeltaRV}) is therefore $l_\mathrm{max} + 5$.

@@

While the R-M signal model is always developed with \texttt{starry} in our analysis, we test multiple approaches to model the effects of the star on these observations.
Over the six hours of the transit, the apparent radial velocity of the star increases by more than 50 m s$^{-1}$. 
In Section \ref{sec:trend} we argue this trend is due to stellar activity rather than an additional unseen planet.
Regardless of the cause of this signal, in order to accurately measure the obliquity of the transiting planet we must model the underlying stellar behavior as well.
We test several different approaches to this problem to ensure our results are not sensitive to our assumptions about the star.
We use low-order polynomials, up to the third degree, to fit the relatively long-term variability during the night of the transit. 
We also build a stellar activity model by fitting a function that is a linear sum of the observed $H\alpha$ and calcium $S_{HK}$ measurements during the night, which are correlated with the observed RV.
Finally, we also fit \texttt{starry} models of a single starspot in time. The spot is modeled as a $l_\mathrm{max} = 4$ spherical harmonic expansion of a Gaussian darker region on the surface of the star; we allow the spot's size, contrast, and location to vary as we fit both signals.

@@

We test multiple likelihood functions with the same form.
First, we assume each data point is drawn from a mixture model which is the sum of two Gaussian functions, each with a different variance:

\begin{equation}
\begin{split}
\mathcal{L} = \prod_i \bigg[ & \frac{q}{\sqrt{2\pi (\sigma^2 + s_1^2)}} \exp\bigg(-\frac{(y_{p,i} + y_{s,i}-V_i)^2}{2(\sigma^2 + s_1^2)}\bigg)  \\+ & \frac{1-q}{\sqrt{2\pi s_2^2)}} \exp\bigg(-\frac{(\alpha y_{p,i} + y_{s,i}-V_i)^2}{2(s_2^2)}\bigg) \bigg],
\end{split}
\end{equation}
Under this model, the $i$th measurement $V_i$ is compared to our model $y_i$. The model is further subdivided into a contribution from the changing background RV baseline due to the star, $y_{s,i}$ and from the planetary signal we are trying to understand, $y_{p,i}$. In the definition of this likelihood function, each data point has some probability $q$ of being a ``good'' data point drawn from a relatively narrow distribution. This distribution is a combination of the PFS pipeline uncertainty, $\sigma$, and an additional jitter term, $s_1$, added in quadrature. 
Each data point also has a probability $1-q$ of being drawn from a much broader distribution, with a separate jitter term, $S_2$, to account for the possibility of stellar effects that significantly affect only a small number of data points. We allow the width of both Gaussian distributions to vary in our fitting procedure. 
Additionally, $q$ is a free parameter: if the data were well-modeled by a single Gaussian, we would find the posterior distribution on $q$ to be consistent with 1.0.
Forcing $q=1$ would thus fit the data using only a single Gaussian.

@@

We only offset the broader Gaussian. Each starspot is unique, and each draw from our posterior tests only a single $\alpha$ value, although disparate starspots with different sizes and different contrast ratios will lead to a different amount of an expected shift from cadence to cadence. 
This uncertainty is manifested through the broader of the two Gaussian terms.
Plainly, this model assumes $q$ is the probability that an observation is not significantly affected by stellar activity. There is then a $1-q$ probability the observation is significantly affected by stellar activity, in which case the magnitude of the effects are relatively uncertain but likely to skew the observation towards smaller absolute values by some factor $\alpha$.
We note we also tested models where $\alpha$ applied to both Gaussian terms, finding this change did not significantly affect our results.

@@

We apply a quadratic limb darkening model for the star, following the prescription of \citet{Kipping13b}. We use uniform priors in both parameters subject to the constraints described in that paper, which enable uninformative sampling of both parameters.

We fit these models to evaluate the posterior distribution using the \texttt{emcee} package of \citet{Foreman-Mackey12}, an implementation of the affine-invariant ensemble sampler of \citet{Goodman10}.
For all explorations using the cubic polynomial out-of-transit model, we initialize 500 walkers; for all other runs we initialize 600 walkers. 
We run each for 3,000 steps, removing the first 2,000 as burn-in and considering the final 1,000 steps in our final analysis. 
We verify that our chains have converged following the method of \citet{Geweke92} and through visual inspection.

Here, we assume the planet orbit is circular. 
Given only information about the transit and assuming the orbital velocity of the planet does not change significantly during the transit there is a degeneracy between the eccentricity and the stellar radius. 
We choose to force circular orbits and fit the orbital separation $a/$\rstar\ as a free parameter. The opposite approach would be equally valid and produce similar results.
We include uniform priors on all parameters except for the projected rotational velocity \vsini\ and the radius ratio $R_p$/\rstar.
For both parameters, \citet{Benatti19} and \citet{Newton19} provide discrepant results; we conservatively apply Gaussian priors with means of 18.3 km s$^{-1}$ and 0.057 and standard deviations of 1.8 km s$^{-1}$ and 0.003 for \vsini\ and $R_p$/\rstar, respectively.
While these papers do predict times of transit and impact parameters, we apply uniform priors on each of these parameters as well to allow the possibility that dynamical interactions have affected the transit timing and impact parameter from the \tess\ epochs to the present day.
When applying a spot model directly through \texttt{starry}, we apply uniform priors on the spot latitude, longitude, contrast ratio, and logarithm of the spot size.

@@

Our results are given in Table \ref{tab:results}. 
The maximum likelihood of the stellar activity indicator model is significantly lower than that of the other models. 
The maximum likelihood $\log \mathcal{L}_\textrm{max}$ differs by only 0.1 between the \texttt{starry} and polynomial models, but is substantially lower for the third model. 
The difference in likelihood corresponds to a Bayes' factor of $\approx 10^{-7}$ when comparing the stellar activity indicator model to the \texttt{starry} model.
In a statistical sense, our starspot model provides approximately an equally valid fit to the data as our simple polynomial model; both provide a much more plausible fit than the stellar activity indicator model.
We note we also tested models using only one of the two stellar activity indicators, but these performed worse than models using both indicators.

All models provide broadly consistent results on the projected obliquity. Considering the two families of most plausible models, the median of the projected obliquity posterior varies from 7 to 14 degrees depending on the specific choice of model used.
All statistical uncertainties range from 11 to 15 degrees.
Using our general polynomial fit, allowing both $q$ and $\alpha$ to vary, we infer a projected obliquity of $14 \pm 11$ degrees.
Likewise, with our \texttt{starry} model we infer a projected obliquity of $12 \pm 13$ degrees.
As this model provides the highest likelihood fit to the data we choose this set of values as most representative of our knowledge of the obliquity of the system, but we emphasize that the particular choice of model or likelihood function does not appear to significantly affect the inferred obliquity.

\citet{Cegla16} provide a relation between the projected obliquity and the true three-dimensional obliquity if the stellar inclination is known. 
The stellar inclination can be inferred from a measurement of the stellar rotation period, \vsini, and stellar radius. 
While there are perhaps significant model uncertainties on the stellar radius, \citet{Newton19} find the stellar inclination consistent with 90 degrees and greater than 70 degrees at 95\% confidence. Assuming these values represent a posterior distribution centered on 90 degrees and with a standard deviation of 10 degrees, we then find the true obliquity between the spin of the host star and orbit of the planet to be less than 16 (27) degrees at $1\sigma$ ($2\sigma$).
